---
title: "additional-analysis"
output: html_document
date: "2025-08-28"
---

```{r}
# Authors: Ian McElveen
# Creation date: 8/28/2025

# LAST UPDATED: 9/2/2025

# Libraries:
library(tidyverse)
```

## 90% Max Opportunities for Starters/key players

The goal is to do the same analysis of reaching 90% maximums, but for starters/key players. Our previous analysis showed that during the in-season, when games are being played, we have a very low count of reaching 90% of a player's all-time maximum velocity. We hypothesize that the reason that counts were lower during the in-season is because less players get opportunities in practices and especially in games. Therefore the in-season dip in counts is because less players have opportunities to have a 90% count, not because of a change in training methods.


```{r}
# Load in Catapult data set
catapult <- read_csv("./Datasets/Catapult Session - Outdoor FB.csv")
```

```{r}
# GAME 1: vs. NDSU

# Filter data set just to include the first game so we can explore easier
ndsu <- catapult |>
  filter(Session.Date == "8/29/2024") |>
  filter(Period.Name %in% c("1ST QUARTER", "2ND QUARTER", "3RD QUARTER", "4TH QUARTER")) |>
  select(anon_id, Date, Primary.Position, Activity, Period.Name, Total.Distance, Total.Acceleration.Efforts, Hit.90.Percent.Max, Hit.Max.Velocity., Maximum.Velocity, All.Time.Max.Velocity)


# Calculate total distance of game for each player
ndsu_totals <- ndsu |>
  group_by(anon_id, Date) |>
  summarise(
    Total.Distance.Game = sum(Total.Distance, na.rm = TRUE),
    Hit.90.Percent.Max = first(Hit.90.Percent.Max),
    .groups = "drop"
  )


# Plot 
ggplot(ndsu_totals, aes(x = reorder(anon_id, Total.Distance.Game), y = Total.Distance.Game, color = Hit.90.Percent.Max)) +
  geom_point() +
  scale_color_manual(values = c("Yes" = "green", "No" = "red")) +
  labs(title = "Hit 90 Percent Max (Yes/No) vs. Total Game Distance - NDSU",
       x = "anon_id", y = "Total Game Distance") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

# T-test between Hit.90 groups (no and yes)
t.test(Total.Distance.Game ~ Hit.90.Percent.Max, data = ndsu_totals)

# No group has mean of 2702 total game distance
# Yes group has mean of 3461 total game distance
# Sig at 0.1 level


# Logistic Regression:

ndsu_totals <- ndsu_totals |>
  mutate(Hit90 = ifelse(Hit.90.Percent.Max == "Yes", 1, 0))

model <- glm(Hit90 ~ Total.Distance.Game, data = ndsu_totals, family = binomial)
summary(model)

exp(coef(model))

```
For the NDSU game, it doesn't seem like more opportunity (greater total game distance) leads to hitting 90 percent of max.




```{r}
# Plots for each game


# UCF game doesn't have quarter names so we have to retrieve this game separetly
ucf <- catapult |>
  filter(Session.Date == "9/28/2024") |>
  select(anon_id, Session.Date, Primary.Position, Activity, Period.Name,
         Total.Distance, Total.Acceleration.Efforts, Hit.90.Percent.Max,
         Hit.Max.Velocity., Maximum.Velocity, All.Time.Max.Velocity) |>
  group_by(Session.Date, anon_id) |>
  summarise(
    Total.Distance.Game = sum(Total.Distance, na.rm = TRUE),
    Hit.90.Percent.Max = first(Hit.90.Percent.Max),
    .groups = "drop"
  )

# Make sure UCF dates are Date objects
ucf <- ucf %>%
  mutate(Session.Date = as.Date(Session.Date, format = "%m/%d/%Y"))  

# Define game dates you care about
game_dates <- as.Date(c("8/29/2024","9/7/2024","9/14/2024","9/21/2024",
                        "9/28/2024","10/12/2024","10/19/2024","10/26/2024",
                        "11/9/2024","11/16/2024","11/23/2024","11/29/2024"),
                      format = "%m/%d/%Y")

# Retrieve data for rest of games
games_data <- catapult |>
  mutate(Session.Date = as.Date(Session.Date, format = "%m/%d/%Y")) |>  # ensure dates
  filter(Session.Date %in% game_dates) |>
  filter(Period.Name %in% c("1ST QUARTER", "2ND QUARTER", "3RD QUARTER", "4TH QUARTER",
                            "1st Quarter", "4th Quarter", "2nd Quarter", "3rd Quarter",
                            "1st QUARTER", "2nd QUARTER", "3rd QUARTER", "4th QUARTER")) |>
  group_by(Session.Date, anon_id) |>
  summarise(
    Total.Distance.Game = sum(Total.Distance, na.rm = TRUE),
    Hit.90.Percent.Max = first(Hit.90.Percent.Max),
    .groups = "drop"
  ) |>
  arrange(Session.Date)

# Ensure games_data has proper Date
games_data <- games_data %>%
  mutate(Session.Date = as.Date(Session.Date))

# Combine with UCF game data
games_data <- bind_rows(games_data, ucf)

# Build schedule using actual dates from games_data
unique_dates <- sort(unique(games_data$Session.Date))

schedule <- tibble(
  Session.Date = unique_dates,
  opponent = c(
    "North Dakota State", "Nebraska", "Colorado State",
    "Baylor", "UCF", "Kansas State",
    "Arizona", "Cincinnati", "Texas Tech",
    "Utah", "Kansas", "Oklahoma State"
  )
)

# Join opponent names into games_data
games_data <- games_data %>%
  left_join(schedule, by = "Session.Date")





# Loop over dates
for (d in unique_dates) {
  
  # Ensure d is Date
  d <- as.Date(d)
  
  # Filter data for this date
  df_sub <- games_data %>% filter(Session.Date == d)
  
  # Get opponent name
  opponent_name <- unique(df_sub$opponent)
  
  # Create plot
  p <- ggplot(df_sub, aes(x = reorder(anon_id, Total.Distance.Game),
                          y = Total.Distance.Game,
                          color = Hit.90.Percent.Max)) +
    geom_point() +
    scale_color_manual(values = c("Yes" = "green", "No" = "red")) +
    labs(title = paste0("Hit 90% Max vs. Total Game Distance - ", opponent_name),
         x = "anon_id", y = "Total Game Distance") +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
  print(p)
}
```

Calculate total game distance over all games for each player.

```{r}
# Plots and tests for all games together


# Calculate total distance for all games and if reached 90% max for any game for all players
player_totals <- games_data |>
  group_by(anon_id) |>
  summarise(
    total_distance_all_games = sum(Total.Distance.Game, na.rm = TRUE),
            reached_90_max = if_else(any(Hit.90.Percent.Max == "Yes"), "Yes", "No")
    ) |>
  arrange(desc(total_distance_all_games))

# Plot
ggplot(player_totals, aes(x = reorder(anon_id, total_distance_all_games), y = total_distance_all_games, color = reached_90_max)) +
  geom_point() +
  scale_color_manual(values = c("Yes" = "green", "No" = "red")) +
  labs(title = "Hit 90% Max vs. Total Distance (All Games)",
       x = "anon_id", y = "Total Distance") +
  theme_classic() +
    theme(axis.text.x = element_text(angle = 90))


ggplot(player_totals, aes(x = reached_90_max, y = total_distance_all_games, fill = reached_90_max)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.1, alpha = 0.5) +
  scale_fill_manual(values = c("Yes" = "green", "No" = "red")) +
  labs(title = "Total Distance vs. Reached 90% Max",
       x = "Reached 90% Max", y = "Total Distance (All Games)") +
  theme_classic()

# T-test
t.test(total_distance_all_games ~ reached_90_max, data = player_totals)


# Convert reached_90_max to 0/1
player_totals <- player_totals |>
  mutate(reached_90_max_num = if_else(reached_90_max == "Yes", 1, 0))

# Model
model <- glm(reached_90_max_num ~ total_distance_all_games,
                   data = player_totals,
                   family = binomial)

summary(model)
```
  Our t-test showed a statically significant difference in total_distance_all_games between the groups.
Players who reached 90% of their maximum velocity during games had significantly greater total game distance than those who did not. The difference is large (~19,650) and statistically reliable (p < 0.001).

  Our linear regression model shows that players who run more distance across games are significantly more likely to reach 90% of their max velocity. The relationship is small per yard, but scales up meaningfully across thousands of yards.
  
For games, more opportunity (yards) equals higher chance of hitting 90% of max.




#### Now we will explore what happens to our 90% count over time graph.

Our 90% count over time graph in our original analysis showed that in weeks during the season, we had the lowest 90% counts. 
We believe this may be because less players have opportunity during those weeks. In-season, practices and games opportunities are mostly taken by the top guys in each position. Therefore, other players would not have the opportunity to hit 90% of their max, leading the in-season to have low counts. Less players getting opportunity = less players hitting 90% of their max.

We will explore this by filtering out players who had less than 10,000 yards of total distance across all of the games in the 2024-25 season. Then we will remake the 90% count over time graph, this time it will only include those top players that get game and practice opportunity.

```{r}

# Get player ids of athletes who did not reach at least 10,000 yards of total distance throughout the season.
high_distance_players <- player_totals |>
  filter(total_distance_all_games > 10000) |>
  pull(anon_id)

# Create data set with those players (don't filter the current data set because we need different data than what our current data set includes)
catapult <- catapult |>
  mutate(Date = as.Date(Date, "%m/%d/%Y"))

top_players <- catapult |>
  filter(anon_id %in% high_distance_players) |>
  filter(Date >= as.Date("2024-06-30") & Date <= as.Date("2025-07-01")) |> # Filter for training season
  filter(Hit.90.Percent.Max == "Yes") |>
  distinct(anon_id, Date, Primary.Position) |>
  group_by(anon_id, Primary.Position) |>
  summarise(times_hit_90 = n(), .groups = "drop") |>
  mutate(
    Position_Group = case_when(
      Primary.Position %in% c("QB", "LB", "TE", "RB") ~ "COMBO",
      Primary.Position %in% c("OL", "DL", "DE") ~ "BIGS",
      Primary.Position %in% c("WR", "DB", "DB, WR") ~ "SKILL",
      TRUE ~ "OTHER"
    ))


# Plot of all players' frequencies
ggplot(top_players, aes(x = anon_id, y = times_hit_90)) +
  geom_bar(stat = "identity", fill = "#CFB87C") +
  geom_hline(yintercept = mean(top_players$times_hit_90), 
             linetype = "dashed", color = "#565A5C", linewidth = 0.5) +
  labs(title = "Player Counts for Achieving ≥ 90% of Maximum Velocity During 2024–25 Season", 
       subtitle = paste("Team Average:", round(mean(top_players$times_hit_90), 1)),
       x = "Athlete ID", y = "Times ≥ 90%") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90))
```
Team average increased from 11.8 to 13.6. There still are some players with really low counts here.

```{r}
# Plots for each position

# Get the overall team average once from hit_90_counts
team_avg <- mean(top_players$times_hit_90)

# Define  positions to loop through
positions <- unique(top_players$Primary.Position)

# Plotting function:
plot_hit_90_by_position <- function(pos) {
  position_data <- top_players %>%
    filter(Primary.Position == pos)
  
  pos_avg <- mean(position_data$times_hit_90)
  
  ggplot(position_data, aes(x = anon_id, y = times_hit_90)) +
    geom_bar(stat = "identity", fill = "#CFB87C") +
    geom_text(aes(label = times_hit_90), vjust = -0.5, size = 3.5) +
    geom_hline(yintercept = pos_avg, linetype = "dashed", color = "#565A5C") +
    annotate("text", x = 1, y = pos_avg + 0.5, 
             label = "Pos Avg", color = "#565A5C", size = 3, hjust = 0) +
    geom_hline(yintercept = team_avg, linetype = "dashed", color = "#000000") +
    annotate("text", x = 1, y = team_avg + 0.5, 
             label = "Team Avg", color = "#000000", size = 3, hjust = 0) +
    labs(title = paste("Times ≥90% Max Velocity –", pos),
         subtitle = paste0("Position Avg: ", round(pos_avg, 1),
                           " | Team Avg: ", round(team_avg, 1)),
         y = "Count",
         x = "anon_id") +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Loop through each position and print the plot
plots_by_position <- lapply(positions, function(pos) {
  print(plot_hit_90_by_position(pos))
})
```
DL: was 10.8, now 11.2
OL: was 6.8, now 7.5
WR: was 15.7, now 23.1
DB: was 18.4, now 20.5
LB: was 8.9, now 7.3
RB: was 10.5, now 6.7
QB: was 9.8, now 3
TE: was 8.4, now 7.5

```{r}
# Make facet plot for position groups so that it fits on the presentation slides better.

# Calculate averages for each group and the team
group_avgs <- top_players %>%
  group_by(Position_Group) %>%
  summarise(group_avg = mean(times_hit_90))

team_avg <- mean(top_players$times_hit_90)

# Join group averages back to the main data
plot_data <- top_players %>%
  left_join(group_avgs, by = "Position_Group")

# Create label data with formatted strings
avg_labels <- plot_data %>%
  group_by(Position_Group) %>%
  summarise(
    group_avg = unique(group_avg),
    team_avg = team_avg,
    label_x = 1,
    group_label_y = group_avg + 1,
    team_label_y = team_avg + 1,
    group_label = round(group_avg, 1),
    team_label = round(team_avg, 1),
    .groups = "drop"
  )

# Plot
ggplot(plot_data, aes(x = anon_id, y = times_hit_90)) +
  geom_bar(stat = "identity", fill = "#CFB87C") +
  geom_hline(aes(yintercept = group_avg), linetype = "dashed", color = "#000000") +
  geom_hline(yintercept = team_avg, linetype = "dotted", color = "#565A5C") +
  geom_text(data = avg_labels, aes(x = label_x, y = group_label_y, label = group_label),
            inherit.aes = FALSE, hjust = 0, size = 3, color = "#000000") +
  geom_text(data = avg_labels, aes(x = label_x, y = team_label_y, label = team_label),
            inherit.aes = FALSE, hjust = 0, size = 3, color = "#565A5C") +
  facet_wrap(~ Position_Group, scales = "free_x") +
  labs(title = "Times Reached ≥90% Max Velocity by Position Group",
       y = "Count", x = "anon_id") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
Bigs: was 8.9, now 9.9
Combo: was 9.3, now 6.5
Skill: was 17.1, 21.6

```{r}
# >90% counts over the course of the season

# Trying to find out when are players reaching above 90% the most

# Create data set that has the count of ≥90% of each day
daily_90_counts <- catapult |>
  filter(anon_id %in% high_distance_players) |>
  filter(Date >= as.Date("2024-06-30") & Date <= as.Date("2025-07-01")) |>
  filter(Hit.90.Percent.Max == "Yes") |>
  distinct(anon_id, Date) |>
  group_by(Date) |>
  summarise(daily_hits = n())

# Plot
ggplot(daily_90_counts, aes(x = Date, y = daily_hits)) +
  geom_line(color = "#CFB87C", linewidth = 1) +
  geom_smooth(method = "loess", se = FALSE, color = "#565A5C", linetype = "dashed") +
  labs(title = "Daily Count of Players Reaching ≥ 90% of Max Velocity",
       subtitle = "Over the 2024–25 Training Season",
       x = "Date", y = "Times reached ≥ 90%") +
  theme_classic()
```

```{r}
# ≥90% counts for each week instead of each day. Will be a little less noisy than the daily
weekly_90_counts <- catapult |>
  filter(anon_id %in% high_distance_players) |>
  filter(Date >= as.Date("2024-06-30") & Date <= as.Date("2025-07-01")) |>
  filter(Hit.90.Percent.Max == "Yes") |>
  distinct(anon_id, Date) |>
  mutate(week = floor_date(Date, unit = "week")) |>
  group_by(week) |>
  summarise(weekly_hits = n())

mean_val <- round(mean(weekly_90_counts$weekly_hits), 1)

ggplot(weekly_90_counts, aes(x = week, y = weekly_hits)) +
   # Add shaded area for season
  annotate("rect",
           xmin = as.Date("2024-08-29"),
           xmax = as.Date("2024-12-28"),
           ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "gray") +
  geom_line(color = "#CFB87C", linewidth = 1) +
  geom_point(color = "#000000") +
  geom_smooth(method = "loess", se = FALSE, color = "#565A5C", linetype = "dashed", linewidth = 0.6) +
  # Monthly x-axis
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b\n%Y"
  ) +
  labs(title = "Weekly Count of Players Reaching ≥ 90% of Max Velocity",
       subtitle = paste0("Week Average: ", mean_val),
       x = "Week", y = "Times reached ≥ 90%") +
  theme_classic()
```

* Note: This graph only includes players who reached at least 10000 total game distance yards during the 2024-25 season.

After changing which players were involved in our analysis, we see a much different trend of 90% counts over the season. We see a steady decrease as the season goes on instead of a U shape that we saw before.




# Does this change our HSI analysis?

Since our athletes have changed, lets look to see if this changes the results we found in our original analysis looking at hamstring injuries.

```{r}
# Load in incident data set
incident <- read_csv("./Datasets/Incident Report FB.csv")
```

```{r}
# Filter data set to only include hamstring injuries

hamstring_injuries <- incident |>
  mutate(Date = as.Date(Date, format = "%m/%d/%Y")) |>
  filter(Date >= as.Date("2024-06-30") & Date <= as.Date("2025-07-01")) |>
  filter(OSICS14.Code %in% c("TR1", "TM1", "TMB", "KTB", "KTH", "TMY", "TMS")) |> 
  filter(anon_id %in% high_distance_players) |>
  distinct(anon_id, Date, .keep_all = TRUE) |>
  mutate(week = floor_date(Date, unit = "week", week_start = 1)) |>
  select(anon_id, Date, week, Position, OSICS.Diagnosis, OSICS14.Code)
```

```{r}
# Overlay injuries onto time graph
ggplot(weekly_90_counts, aes(x = week, y = weekly_hits)) +
  # Add shaded area for season
  annotate("rect",
           xmin = as.Date("2024-08-29"),
           xmax = as.Date("2024-12-28"),
           ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "gray") +
  geom_line(color = "#CFB87C", linewidth = 1) +
  geom_point(color = "#000000") +
  geom_smooth(method = "loess", se = FALSE, color = "#565A5C", linetype = "dashed", linewidth = 0.6) +
  # Injury markers
  geom_vline(data = hamstring_injuries, aes(xintercept = week),
             color = "red", linetype = "dashed", linewidth = 0.6) +
  # Plot from Q1
  geom_line(color = "#CFB87C", linewidth = 1) +
  geom_point(color = "#000000") +
  geom_smooth(method = "loess", se = FALSE, color = "#565A5C", linetype = "dashed", linewidth = 0.6) +

  # Monthly x-axis
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b\n%Y"
  ) +

  labs(
    title = "Weekly Count of Players Reaching ≥ 90% of Max Velocity with Injuries",
    subtitle = paste0("Week Average: ", mean_val),
    x = "Month", y = "Times reached ≥ 90%"
  ) +
  theme_classic()

hamstring_weeks <- hamstring_injuries |>
  count(week, name = "injury_count")

ggplot(weekly_90_counts, aes(x = week, y = weekly_hits)) +
  # Add shaded area for season
  annotate("rect",
           xmin = as.Date("2024-08-29"),
           xmax = as.Date("2024-12-28"),
           ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "gray") +
  geom_line(color = "#CFB87C", linewidth = 1) +
  geom_point(color = "#000000") +
  geom_smooth(method = "loess", se = FALSE, color = "#565A5C", 
              linetype = "dashed", linewidth = 0.6) +
  
  # Single injuries
  geom_vline(data = filter(hamstring_weeks, injury_count == 1),
             aes(xintercept = week),
             color = "red", linetype = "dotted", linewidth = 0.6) +
  
  # Multiple injuries - solid, thicker line
  geom_vline(data = filter(hamstring_weeks, injury_count >= 2),
             aes(xintercept = week),
             color = "red", linetype = "dashed", linewidth = 0.6) +
  
  # Monthly x-axis
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b\n%Y"
  ) +
  
  labs(
    title = "Weekly Count of Players Reaching ≥ 90% of Max Velocity with Injuries",
    subtitle = paste0("Week Average: ", mean_val,
                     "\nDotted red: 1 injury | Dashed red: 2 injuries"),
    x = "Month", 
    y = "Times reached ≥ 90%"
  ) +
  theme_classic()





# First, create a combined dataset for injuries with a type variable
injury_type <- bind_rows(
  filter(hamstring_weeks, injury_count == 1) |> mutate(injury_type = "1 injury"),
  filter(hamstring_weeks, injury_count >= 2) |> mutate(injury_type = "2 injuries")
)


ggplot(weekly_90_counts, aes(x = week, y = weekly_hits)) +
  # Add shaded area for season
  annotate("rect",
           xmin = as.Date("2024-08-29"),
           xmax = as.Date("2024-12-28"),
           ymin = -Inf, ymax = Inf,
           alpha = 0.2, fill = "gray") +
  geom_line(color = "#CFB87C", linewidth = 1) +
  geom_point(color = "#000000") +
  geom_smooth(method = "loess", se = FALSE, color = "#565A5C", 
              linetype = "dashed", linewidth = 0.6) +
  
  # Add injury lines with aesthetic mapping for legend
  geom_vline(data = injury_type,
             aes(xintercept = week, linetype = injury_type, color = injury_type),
             linewidth = 0.6, show.legend = TRUE) +
  
  # Customize line styles and colors
  scale_linetype_manual(
    name = "Hamstring Injuries",
    values = c("1 injury" = "dotted", "2 injuries" = "dashed")
  ) +
  scale_color_manual(
    name = "Hamstring Injuries",
    values = c("1 injury" = "red", "2 injuries" = "red")
  ) +
  
  # Monthly x-axis
  scale_x_date(
    date_breaks = "1 month",
    date_labels = "%b\n%Y"
  ) +
  
  labs(
    title = "Weekly Count of Players Reaching ≥ 90% of Max Velocity with Injuries",
    subtitle = paste0("Week Average: ", mean_val),
    x = "Month", 
    y = "Times reached ≥ 90%"
  ) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    legend.title = element_text()
  )
```

```{r}
# Daily

ggplot(daily_90_counts, aes(x = Date, y = daily_hits)) +
  geom_line(color = "#CFB87C", linewidth = 1) +
  geom_smooth(method = "loess", se = FALSE, color = "#565A5C", linetype = "dashed") +
  geom_vline(data = hamstring_injuries, aes(xintercept = Date),
             color = "red", linetype = "dashed", linewidth = 0.6) +
  labs(title = "Daily Count of Players Reaching ≥ 90% of Max Velocity",
       subtitle = "Over the 2024–25 Training Season",
       x = "Date", y = "Times reached ≥ 90%") +
  theme_classic()
```

```{r}
# Create count of 90 percent efforts for each player and week
weekly_player_90_counts <- catapult |>
  mutate(Date = as.Date(Date, format = "%m/%d/%Y")) |>
  filter(Date >= as.Date("2024-06-30") & Date <= as.Date("2025-07-01")) |>
  filter(anon_id %in% high_distance_players) |>
  distinct(anon_id, Date, .keep_all = TRUE) |>
  select(anon_id, Date, Primary.Position, Activity, Hit.90.Percent.Max) |>
  mutate(week = floor_date(Date, unit = "week", week_start = 1)) |>
  group_by(anon_id, week) |>
  summarise(
    hits_90 = sum(Hit.90.Percent.Max == "Yes", na.rm = TRUE),
    total_sessions = n(),
    .groups = "drop"
  ) |>
  mutate(lastweek_hits_90 = lag(hits_90, 1),
         last2weeks_hits_90 = lag(hits_90, 2),
         last3weeks_hits_90 = lag(hits_90, 3),
         
         sum_last = hits_90 + lastweek_hits_90,
         sum_last2 = hits_90 + lastweek_hits_90 + last2weeks_hits_90,
         sum_last3 = hits_90 + lastweek_hits_90 + last2weeks_hits_90 + last3weeks_hits_90,
  )

# Join with hamstring_injuries data set
weekly_with_injuries <- weekly_player_90_counts |>
  left_join(hamstring_injuries, by = c("anon_id", "week")) |>
  mutate(injury = ifelse(is.na(OSICS14.Code), 0, 1))

```

```{r}
# Make bar plot of injuries by 90 counts and injury rates

injury_counts <- weekly_with_injuries |>
  group_by(hits_90) |>
  summarise(
    n_injuries = sum(injury),
    total_weeks = n(),
    .groups = "drop"
  ) |>
  mutate(injury_rate = round((n_injuries / total_weeks) * 100, 1))

# Last 3 weeks
injury_counts_3_weeks <- weekly_with_injuries |>
  group_by(sum_last3) |>
  summarise(
    n_injuries = sum(injury),
    total_weeks = n(),
    .groups = "drop"
  ) |>
  mutate(injury_rate = round((n_injuries / total_weeks) * 100, 1))

# Number of Injuries
ggplot(injury_counts, aes(x = hits_90, y = n_injuries)) +
  geom_bar(stat = "identity", fill = "#CFB87C") +
  geom_text(aes(label = n_injuries),
            vjust = -0.2,
            size = 3.5) +
  labs(
    title = "Number of Hamstring Injuries by ≥90% Counts",
    subtitle = "Week of an Injury",
    x = "≥90% Counts",
    y = "Number of Injuries"
  ) +
  theme_classic()

# Injury Rate
ggplot(injury_counts, aes(x = hits_90, y = injury_rate)) +
  geom_bar(stat = "identity", fill = "#CFB87C") +
  geom_text(aes(label = injury_rate),
            vjust = -0.2,
            size = 3.5) +
  labs(
    title = "Injury Rate by ≥90% Counts",
    subtitle = "Week of an Injury",
    x = "≥90% Counts",
    y = "Injury Rate (%)"
  ) +
  theme_classic()


# Count sum over last 3 weeks before an injury
# Number of Injuries
ggplot(injury_counts_3_weeks, aes(x = sum_last3, y = n_injuries)) +
  geom_bar(stat = "identity", fill = "#CFB87C") +
  geom_text(aes(label = n_injuries),
            vjust = -0.2,
            size = 3.5) +
  scale_x_continuous(breaks = seq(min(injury_counts_3_weeks$sum_last3, na.rm = TRUE),
                                  max(injury_counts_3_weeks$sum_last3, na.rm = TRUE), 1)) +
  labs(
    title = "Number of Hamstring Injuries by ≥90% Counts",
    subtitle = "Total over last 3 weeks before injury",
    x = "≥90% Counts",
    y = "Number of Injuries"
  ) +
  theme_classic()

# Injury Rate
ggplot(injury_counts_3_weeks, aes(x = sum_last3, y = injury_rate)) +
  geom_bar(stat = "identity", fill = "#CFB87C") +
  geom_text(aes(label = injury_rate),
            vjust = -0.2,
            size = 3.5) +
  scale_x_continuous(breaks = seq(min(injury_counts_3_weeks$sum_last3, na.rm = TRUE),
                                  max(injury_counts_3_weeks$sum_last3, na.rm = TRUE), 1)) +
  labs(
    title = "Injury Rate by ≥90% Counts",
    subtitle = "Week of an Injury",
    x = "≥90% Counts",
    y = "Injury Rate (%)"
  ) +
  theme_classic()
```

```{r}
# Test if injury rates are significant

# Create the data frame
injury_data <- data.frame(
  sum_last3 = c(0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13),
  n_injuries = c(2, 7, 1, 3, 0, 1, 1, 0, 0, 0, 0, 0, 0, 0),
  total_weeks = c(506, 339, 236, 124, 108, 47, 41, 22, 13, 10, 7, 3, 3, 1)
)

# Fit the Poisson regression model
# This models the injury count based on the number of sprints, accounting for different weeks of exposure
model <- glm(n_injuries ~ sum_last3, 
             family = poisson, 
             data = injury_data, 
             offset = log(total_weeks))

# Get a summary of the results
summary(model)
```
