---
title: "Data Analysis Notebook"
output: html_notebook
---

```{r}
#Authors: [INSERT NAMES HERE]
#Author Date: 
#Purpose: The purpose of this notebook is to house all data set transformation, cleansing, visualization, statistical analysis, and note-taking for the 2025 CU Athletic Department Sports Science Internship Program

#LAST UPDATED: 

#Including helpful libraries
library(tidyverse)
library(readxl)
library(aod)
library(gt)
```


# Data Cleaning
```{r Loading in the data sets}
#loading in the Catapult data to look at sprinting values
Catapult_Session <- read_csv("data-sets/data-sets-uncompressed/data-sets-compressed/Running Imbalance and Speed/Catapult Session - Outdoor FB.csv")

#loading in the Historical Running data to look at running imbalance values
Historical_Running <- read_csv("data-sets/data-sets-uncompressed/data-sets-compressed/Running Imbalance and Speed/Compiled Historical Running Imbalance FB.csv")

#loading in the Incident Report to look at HSIs
Incident_Report <- read_csv("data-sets/data-sets-uncompressed/data-sets-compressed/Running Imbalance and Speed/Incident Report FB IDs.csv")

```

```{r Cleaning Catapult_Session}
Catapult_Session_clean <- Catapult_Session %>%
  #putting the date as a date class
  mutate(Date = as.Date(Date, "%m/%d/%Y")) %>%
  #only selecting important columns for this analysis
  select(anon_id, Date, Age, Primary.Position, Total.Distance, Period.Name, Total.Duration..min., Velocity.Band.1.Total.Distance, Velocity.Band.2.Total.Distance, Velocity.Band.3.Total.Distance, Velocity.Band.4.Total.Distance, Velocity.Band.5.Total.Distance, Velocity.Band.6.Total.Distance, Velocity.Band.7.Total.Distance, Velocity.Band.8.Total.Distance, Maximum.Velocity, Average.Velocity, Hit.90.Percent.Max, Date.of.Last.90.Effort, Days.Since.Last.90.Effort) %>%
  #calculating each player's maximum velocity
  group_by(anon_id) %>%
  mutate(Player.Max.Velocity = max(na.omit(Maximum.Velocity))) %>%
  ungroup() %>%
  #only selecting data from January 1, 2024 and on
  filter(Date >= "2024-01-01")

head(Catapult_Session_clean)
```

```{r Cleaning Historical_Running}
Historical_Running_clean <- Historical_Running %>%
  #taking out rows that don't have data
  filter(Running.Imbalance != "n/a") %>%
  #putting running imbalance as a number and converting the date to a date class
  mutate(Running.Imbalance = as.numeric(Running.Imbalance),
         Date = as.Date(Date, "%m/%d/%Y")) %>%
  #only using data from January 1, 2024 and on
  filter(Date >= "2024-01-01") %>%
  mutate(X=1:4063) %>%
  #making days January 1, 2024
  group_by(anon_id) %>%
  mutate(Days.Since.Start = as.numeric(Date - min(Date))) %>%
  ungroup()

head(Historical_Running_clean)
```

```{r Cleaning Incident_Report}
Incident_Report_clean <- Incident_Report %>%
  #filtering for only hamstring injuries
  filter(OSICS14.Code == "TM1",
         Status != "Full Go")  %>%
  #getting the date of the injury as a date class
  mutate(Date = as.Date(Date, "%m/%d/%Y"),
         Date.of.Injury = as.Date(Date.of.Injury...Onset.of.symptoms, "%m/%d/%Y"),
         Examination.Date = as.Date(Examination.Date, "%m/%d/%Y")) %>%
  #only selecting relevant columns for this analysis
  select(anon_id, Position, Date, Date.of.Injury, Time.of.Injury, Side, OSICS.Injury.Diagnosis, Coach.s.Diagnosis, Recurrence.of.Injury, Choose.Season, Onset.of.Symptoms, Injury.Prognosis, General.Mechanism, Specific.Mechanism, Injured.While., Type.of.Event, Season., Status, Days.in.Status) %>%
  group_by(anon_id, Date.of.Injury) %>%
  mutate(Days.Out = sum(Days.in.Status)) %>%
  ungroup()

head(Incident_Report_clean)
```

```{r Identifying players that have data in both data sets}
#taking the IDs of players who are and aren't injured
all_IDs <- unique(Historical_Running_clean$anon_id)
#taking IDs that were injured and also have running imbalance data
injured_IDs <- intersect(unique(Incident_Report_clean$anon_id), all_IDs)
#taking all players with running imbalance data that don't have an injury
uninjured_IDs <- unique((Historical_Running_clean %>%
  filter(!anon_id %in% injured_IDs))$anon_id)
```

```{r Filtering out players without proper data from all data sets}
#injured players who also have running imbalance data
Incident_Report_clean <- Incident_Report_clean %>%
  filter(anon_id %in% injured_IDs)

#all players that only have running imbalance data or have both running imbalance data and incidence report
Historical_Running_clean <- Historical_Running_clean %>%
  filter(anon_id %in% injured_IDs | anon_id %in% uninjured_IDs)
```

```{r Removing all unimportant objects created during cleaning}
#removing all_IDs vector and uncleaned data sets
remove(all_IDs)
remove(Incident_Report)
remove(Historical_Running)
remove(Catapult_Session)
```

# Section 1: Running Speed

## How often are athletes reaching ≥ 90% maximum velocity throughout a training season?

```{r}
# Bar chart for how often each player reaches ≥ 90% maximum velocity

# Count for how many times each anon_id hit ≥ 90% maximum velocity
hit_90_counts <- Catapult_Session_clean %>%
  filter(Date >= as.Date("2024-06-30") & Date <= as.Date("2025-07-01")) %>%   # Filter for training season
  filter(Hit.90.Percent.Max == "Yes") %>%
  distinct(anon_id,Date) %>%
  group_by(anon_id) %>%
  summarise(times_hit_90 = n())


# Plot of all players' frequencies
ggplot(hit_90_counts, aes(x = anon_id, y = times_hit_90)) +
  geom_bar(stat = "identity", fill = "#CFB87C") +
  geom_hline(yintercept = mean(hit_90_counts$times_hit_90), 
             linetype = "dashed", color = "#565A5C", linewidth = 0.5) +
  labs(title = "Player Counts for Achieving ≥ 90% of Maximum Velocity During 2024–25 Season", 
       subtitle = paste("Team Average:", round(mean(hit_90_counts$times_hit_90), 1)),
       x = "Athlete ID", y = "Times ≥ 90%") +
  theme_classic() +
  theme(axis.text.x = element_text(size = 6, angle = 90))
```
# Delete because we created a function to plot each position
```{r}
hit_90_counts <- Catapult_Session_clean %>%
  filter(Date >= as.Date("2024-06-30") & Date <= as.Date("2025-07-01"),
         Hit.90.Percent.Max == "Yes") %>%
  distinct(anon_id, Date, .keep_all = TRUE) %>%  # Keep position data
  group_by(anon_id, Primary.Position) %>%        # Group by both ID and position
  summarise(times_hit_90 = n(), .groups = "drop")

QBs <- hit_90_counts %>%
  filter(Primary.Position == "QB")

# Calculate the averages first
overall_avg <- mean(hit_90_counts$times_hit_90)
qb_avg <- mean(QBs$times_hit_90)

ggplot(QBs, aes(x=anon_id, y=times_hit_90)) +
  geom_bar(stat="identity", fill = "#CFB87C") + 
  geom_text(aes(label = times_hit_90), 
            vjust = -0.5, 
            size = 3.5) +
  geom_hline(yintercept = overall_avg, 
             linetype = "dashed", color = "#000000") +
  geom_hline(yintercept = qb_avg, 
             linetype = "dashed", color = "#565A5C") +
  annotate("text", x = 1, y = overall_avg + 0.5, 
           label = "Team Avg", color = "#000000", size = 3) +
  annotate("text", x = 1, y = qb_avg + 0.5, 
           label = "QB Avg", color = "#565A5C", size = 3) +
  labs(title = "QB counts for reaching > 90% max velocity", 
       subtitle = paste("QB Average:", qb_avg)) +
  theme_classic()

```
## Maybe delete facet plot, as it is hard to read
```{r}
# Facet Plot

# Recalculate hit_90_counts for all positions
hit_90_counts <- Catapult_Session_clean %>%
  filter(Date >= as.Date("2024-06-30") & Date <= as.Date("2025-07-01"),
         Hit.90.Percent.Max == "Yes") %>%
  distinct(anon_id, Date, .keep_all = TRUE) %>%  # Keep position data
  group_by(anon_id, Primary.Position) %>%        # Group by both ID and position
  summarise(times_hit_90 = n(), .groups = "drop")

# Calculate overall average
overall_avg <- mean(hit_90_counts$times_hit_90)

# Plot faceted bar charts by position
ggplot(hit_90_counts, aes(x = anon_id, y = times_hit_90)) +
  geom_bar(stat = "identity", fill = "#CFB87C") +
  geom_text(aes(label = times_hit_90), vjust = -0.5, size = 3.5) +
  geom_hline(yintercept = overall_avg, linetype = "dashed", color = "#000000") +
  annotate("text", x = 1, y = overall_avg + 0.5, 
           label = "Team Avg", color = "#000000", size = 3, hjust = 0) +
  facet_wrap(~ Primary.Position, scales = "free_x") +
  labs(title = "Counts of >90% Max Velocity by Player and Position",
       y = "Times > 90% Max Velocity",
       x = "Player (anon_id)") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
# Plots for each position

hit_90_counts <- Catapult_Session_clean %>%
  filter(Date >= as.Date("2024-06-30") & Date <= as.Date("2025-07-01"),
         Hit.90.Percent.Max == "Yes") %>%
  distinct(anon_id, Date, .keep_all = TRUE) %>%
  group_by(anon_id, Primary.Position) %>%
  summarise(times_hit_90 = n(), .groups = "drop")

# Get the overall team average once from hit_90_counts
team_avg <- mean(hit_90_counts$times_hit_90)

# Define  positions to loop through
positions <- unique(hit_90_counts$Primary.Position)

# Plotting function:
plot_hit_90_by_position <- function(pos) {
  position_data <- hit_90_counts %>%
    filter(Primary.Position == pos)
  
  pos_avg <- mean(position_data$times_hit_90)
  
  ggplot(position_data, aes(x = anon_id, y = times_hit_90)) +
    geom_bar(stat = "identity", fill = "#CFB87C") +
    geom_text(aes(label = times_hit_90), vjust = -0.5, size = 3.5) +
    geom_hline(yintercept = pos_avg, linetype = "dashed", color = "#565A5C") +
    annotate("text", x = 1, y = pos_avg + 0.5, 
             label = "Pos Avg", color = "#565A5C", size = 3, hjust = 0) +
    geom_hline(yintercept = team_avg, linetype = "dashed", color = "#000000") +
    annotate("text", x = 1, y = team_avg + 0.5, 
             label = "Team Avg", color = "#000000", size = 3, hjust = 0) +
    labs(title = paste("Times ≥90% Max Velocity –", pos),
         subtitle = paste0("Position Avg: ", round(pos_avg, 1),
                           " | Team Avg: ", round(team_avg, 1)),
         y = "Count",
         x = "anon_id") +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Loop through each position and print the plot
plots_by_position <- lapply(positions, function(pos) {
  print(plot_hit_90_by_position(pos))
})

```

```{r}
# Table for average values of each position

position_averages <- hit_90_counts %>%
  group_by(Primary.Position) %>%
  summarise(avg_times_hit_90 = mean(times_hit_90), .groups = "drop") %>%
  arrange(desc(avg_times_hit_90))

position_averages_with_team <- bind_rows(
  tibble(
    Primary.Position = "Team Average",
    avg_times_hit_90 = team_avg
  ),
  position_averages
)

position_averages_with_team %>%
  gt() %>%
  tab_header(
    title = "Average Times >90% Max Velocity by Position and Team"
  )
```

## Should we consider the number of sprinting efforts that athletes are completing?

```{r}
# Create weekly sprint count totals

# Create a week variable
Catapult_Session_clean <- Catapult_Session_clean %>%
  mutate(week = floor_date(Date, unit = "week"))

# Get unique athletes and all weeks in your date range
all_athletes <- unique(Catapult_Session_clean$anon_id)
all_weeks <- seq.Date(as.Date("2024-06-30"), as.Date("2025-07-01"), by = "week")

# Create a full grid of anon_id and weeks (this includes weeks with 0 sprints)
full_grid <- expand_grid(anon_id = all_athletes, week = all_weeks)

# Count sprints per athlete-week
sprint_counts <- Catapult_Session_clean %>%
  filter(Hit.90.Percent.Max == "Yes") %>%
  distinct(anon_id, Date, .keep_all = TRUE) %>%
  group_by(anon_id, week) %>%
  summarise(sprint_count = n(), .groups = "drop")

# Join full grid with actual sprint counts and replace NAs with 0s
weekly_sprint_counts <- full_grid %>%
  left_join(sprint_counts, by = c("anon_id", "week")) %>%
  mutate(sprint_count = replace_na(sprint_count, 0))

# Categorize sprint load groups
weekly_sprint_counts <- weekly_sprint_counts %>%
  mutate(sprint_load_group = case_when(
    sprint_count <= 1 ~ "Low (≤1 sprint)",
    sprint_count %in% 2:3 ~ "Moderate (2–3 sprints)",
    sprint_count %in% 4:5 ~ "High (4–5 sprints)"
  ))
```

```{r}
# Merge weekly sprints and injury data frames

# Create an injury weeks data frame from incident report
# This should indicate which week the injury occurred
injury_weeks <- Incident_Report_clean %>%
  mutate(
    week = floor_date(as.Date(Date.of.Injury), unit = "week")
  ) %>%
  filter(Date.of.Injury >= as.Date("2024-06-30") & Date.of.Injury <= as.Date("2025-07-01")) %>%
  select(anon_id, week) %>%
  distinct() %>%
  mutate(injury = 1)

# Merge with weekly sprints data frame
weekly_sprint_counts <- weekly_sprint_counts %>%
  left_join(injury_weeks, by = c("anon_id", "week")) %>%
  mutate(injury = ifelse(is.na(injury), 0, injury))  # Fill NAs with 0 (no injury)
```

```{r}
# Table to compare injury weeks to non injury weeks
weekly_sprint_counts %>%
  group_by(injury) %>%
  summarise(
    mean_sprints = mean(sprint_count, na.rm = TRUE),
    sd_sprints = sd(sprint_count, na.rm = TRUE),
    n = n()
  )

ggplot(weekly_sprint_counts, aes(x = week, y = sprint_count, color = factor(injury))) +
  geom_line(aes(group = anon_id), alpha = 0.5) +
  geom_point(data = subset(weekly_sprint_counts, injury == 1), size = 2, shape = 21, fill = "red") +
  scale_color_manual(values = c("0" = "grey", "1" = "red"), labels = c("No Injury", "Injury")) +
  labs(title = "Weekly Sprint Counts with Injury Events", x = "Week", y = "Sprint Count", color = "Injury") +
  theme_minimal()

```
No injuries occurred when an athlete had 4 or 5 sprint counts within a week. There are limited data points in these sprint counts.
There are also only three injuries that occur in the moderate range of 2-3 sprint counts
The most injuries occur in the low range of 0-1 sprints.

This suggests that underexposure to sprinting may increase injury risk.

```{r}
# Join to add week number per athlete
weekly_sprint_counts <- weekly_sprint_counts %>%
  group_by(anon_id) %>%
  arrange(week) %>%
  mutate(week_number = row_number()) %>%
  ungroup()

# Identify weeks just before injury
pre_injury_weeks <- weekly_sprint_counts %>%
  group_by(anon_id) %>%
  mutate(
    next_injury = lead(injury),
    is_pre_injury = ifelse(next_injury == 1, 1, 0)
  ) %>%
  ungroup() %>%
  filter(is_pre_injury == 1)

# Compare average sprint counts
pre_injury_weeks %>%
  summarise(mean_pre_injury_sprints = mean(sprint_count, na.rm = TRUE))

weekly_sprint_counts %>%
  filter(injury == 0) %>%
  summarise(mean_non_injury_sprints = mean(sprint_count, na.rm = TRUE))

```
```{r}
ggplot(weekly_sprint_counts, aes(x = factor(injury), y = sprint_count, fill = factor(injury))) +
  geom_boxplot() +
  scale_fill_manual(values = c("0" = "lightblue", "1" = "salmon")) +
  labs(title = "Sprint Count Distribution by Injury Status", x = "Injury", y = "Sprint Count") +
  theme_minimal()


```
## Are relative efforts and bands more advantageous than the absolute bands provided?

## How does sprinting exposure (# of efforts, % max reached) relate to incidence of hamstring injuries?



# Section 2: Running Imbalance

## What is the variation at the team level and at each individual athlete level?

  Each player tends to have a very unique trend in their running imbalance. Looking at how the team varies but also at how each player varies throughout the season, it's hard to make out any pattern that's applicable to most people. The variance of running imbalance varies greatly between each player. Instead we looked at what are the variances between players who were injured and those who were not. Based on three different bootstrapped findings, we can see that the variances between players who were injured and those who were not were both statistically significant.
  For the first bootstrap, we compared the variance of the pooled groups meaning that the variance in running imbalance for players who were injured and those who were not were compared. This resulted in a 90% confidence interval which suggested that the difference in variance between the two groups is between 0.30 and 3.45. This suggests that when looking at the variance of the two groups separately but all the players are pooled together, the variances will most likely be different by factor between 0.30 and 3.45 and the variance for the injured pool will be greater than that of the uninjured pool.
  For the second bootstrap, each player's variance was taken individually. This unpooled approach was taken to see if a player's variance in running imbalance could potentially be related to HSI risk. The bootstrap algorithm in this case took the averaged variances of the bootstrapped sample for each group and compared them. This bootstrap produced a 90% confidence interval for the difference in average variance between the two groups is 0.79 to 1.32. These results suggest that players who sustained a hamstring injury since January 1, 2024 had, on average, a greater variance in their running imbalance by about 1.07. This suggests that there is a relationship between variance in running imbalance and HSI risk. This found increase in variability will be used to address the following questions.
  For the third bootstrap, we wanted to see if there was a difference in the average mean absolute value in running imbalance between players who were and weren't injured. The bootstrapping algorithm for this test calculated the average absolute distance value or each running imbalance measurement and found the average for each sample. This test found that at the 90% significance level, injured players had an average running imbalance absolute value between 0.06 and 0.32 greater than their uninjured counterparts. These values though, when we consider that the range of running imbalance goes from 0 to 100 is small and may be hard to detect when out in the field. 

```{r Looking at team and player variances of running imbalance}
#team variation
Historical_Running_clean %>%
  summarize(Team_Variation = var(Running.Imbalance))

#individual player variation
Historical_Running_clean %>%
  group_by(anon_id) %>%
  summarize(Player_Variation = var(na.omit(Running.Imbalance))) %>%
  ungroup() 
```


```{r Running imbalance measurements for whole team since January 1, 2024}
#calculating mean and variance for team data
team_mean <- mean(Historical_Running_clean$Running.Imbalance)
team_sd <- sd(Historical_Running_clean$Running.Imbalance)

#making scatter plot of team running imbalance data throughout season
ggplot(Historical_Running_clean, aes(Date, Running.Imbalance)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = team_mean, color = "#CFB87C") +
  geom_hline(yintercept = team_mean + team_sd) +
  geom_hline(yintercept = team_mean - team_sd) +
  geom_hline(yintercept = team_mean + (2*team_sd), color = "#A2A4A3") +
  geom_hline(yintercept = team_mean - (2*team_sd), color = "#A2A4A3") +
  labs(title = "Team Running Imbalance Since January 1, 2024", y="Running Imbalance (%)",
       subtitle = "\u03BC = 0.08623412, \u03C3^2 = 14.94215")

#making histogram of team running imbalance data
ggplot(Historical_Running_clean, aes(Running.Imbalance)) +
  geom_histogram() +
  labs(title = "Team Running Imbalance Since January 1, 2024", x="Running Imbalance")
```


```{r Looking at distributions for injured and uninjured separately}
#making histogram for running imbalance of all injured athletes
ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,], aes(Running.Imbalance)) +
  geom_histogram(fill = "#CFB87C", alpha = 0.75) +
  #adding in 95% confidence interval
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,]$Running.Imbalance, 0.025), color = "#CFB87C") +
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,]$Running.Imbalance, 0.975), color = "#CFB87C") +
  xlim(-21,21) +
  labs(title = "Running Imbalance for Players with HSI since January 1, 2024")

#making histogram for running imbalance of all uninjured athletes
ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,], aes(Running.Imbalance)) +
  geom_histogram(fill = "black", alpha = 0.75) +
  #adding in 95% confidence interval
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,]$Running.Imbalance, 0.025)) +
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,]$Running.Imbalance, 0.975)) +
  xlim(-21,21) +
  labs(title = "Running Imbalance for Players without HSI since January 1, 2024")

#Plotting injured and uninjured histograms over top one another
ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,], aes(Running.Imbalance)) +
  geom_histogram(alpha = 0.75) +
  #adding in 95% CI for uninjured players
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,]$Running.Imbalance, 0.025)) +
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,]$Running.Imbalance, 0.975)) +
  geom_histogram(data = Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,], aes(Running.Imbalance), fill = "#CFB87C", alpha = 0.75) +
  #adding in 95% CI for injured players
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,]$Running.Imbalance, 0.025), color = "#CFB87C") +
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,]$Running.Imbalance, 0.975), color = "#CFB87C") +
  xlim(-21,21) +
  labs(title = "Running Imbalance for Players with and without HSI since January 1, 2024")
```


```{r looking at trends for injured and uninjured players since 1-1-2024}
#making scatter plot of running imbalance data for injured players
ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,], aes(Date, Running.Imbalance)) + 
  geom_point(alpha = 0.3) +
  ylim(-21,21) +
  labs(title = "Running Imbalance for Players with HSI since January 1, 2024")

#making scatter plot of running imbalance for uninjured players
ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,], aes(Date, Running.Imbalance)) + 
  geom_point(alpha = 0.3) +
  ylim(-21,21) +
  labs(title = "Running Imbalance for Players without HSI since January 1, 2024")
```


### Testing to see if significantly different
```{r Separating injured and uninjured into two different data sets}
#Splitting up the data sets into injured and uninjured groups
injured_data <- Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,] %>%
  mutate(Player.Absolute.Dist = abs(Running.Imbalance)) %>%
  group_by(anon_id) %>%
  mutate(Player.Variance = var(Running.Imbalance)) %>%
  ungroup()

uninjured_data <- Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,] %>%
  mutate(Player.Absolute.Dist = abs(Running.Imbalance)) %>%
  group_by(anon_id) %>%
  mutate(Player.Variance = var(Running.Imbalance)) %>%
  ungroup()
```


```{r Boostrapping to look at variances of different groups, pooled}
#making a data frame to hold all of the within group variances
group_variances <- data.frame(injured_var = rep(NA,1000),
                uninjured_var = rep(NA,1000),
                diff_in_var = rep(NA, 1000))

#bootstrap for variances, 1000 iterations
for(i in 1:1000){
  #random seed
  set.seed(i) 
  
  #taking samples from each of the data sets, same number of rows, replacement true
  injured_sample <- sample_n(injured_data, replace = TRUE, size = 1672)
  uninjured_sample <- sample_n(uninjured_data, replace=TRUE, size = 2391)
  
  #storing the calculated variances in data frame
  group_variances[i,1] = var(injured_sample$Running.Imbalance)
  group_variances[i,2] = var(uninjured_sample$Running.Imbalance)
  group_variances[i,3] = group_variances[i,1] - group_variances[i,2]
}
```


```{r Plotting results from pooled boostrapped variances}
ggplot(data=group_variances, aes(diff_in_var)) +
  geom_histogram() +
  geom_vline(xintercept = quantile(group_variances$diff_in_var, 0.05), color= "#CFB87C") +
  geom_vline(xintercept = quantile(group_variances$diff_in_var, 0.95), color= "#CFB87C")

ggplot(data=group_variances) +
  geom_histogram(aes(injured_var), alpha = 0.75, fill ="#CFB87C") +
  geom_histogram(aes(uninjured_var), alpha = 0.75)
```


```{r Bootstrapping to look at average player variances}
#making a data frame to hold all of the average player variances between groups
mean_player_variances <- data.frame(injured_var = rep(NA,1000),
                uninjured_var = rep(NA,1000),
                diff_in_var = rep(NA, 1000))

for(i in 1:1000){
  #random seed
  set.seed(i) 
  
  #taking samples from each of the data sets, same number of rows, replacement true
  injured_sample <- sample_n(injured_data, replace = TRUE, size = 1672)
  uninjured_sample <- sample_n(uninjured_data, replace=TRUE, size = 2391)
  
  #storing the calculated variances in data frame
  mean_player_variances[i,1] = mean(na.omit(injured_sample$Player.Variance))
  mean_player_variances[i,2] = mean(na.omit(uninjured_sample$Player.Variance))
  mean_player_variances[i,3] = mean_player_variances[i,1] - mean_player_variances[i,2]
}
```


```{r Plotting results from averaged bootstrapped player variances}
ggplot(data = mean_player_variances, aes(diff_in_var)) +
  geom_histogram() +
  geom_vline(xintercept = quantile(mean_player_variances$diff_in_var, 0.05), color= "#CFB87C") +
  geom_vline(xintercept = quantile(mean_player_variances$diff_in_var, 0.95), color= "#CFB87C") +
  labs(x="Difference in Variance")

ggplot(data=mean_player_variances) +
  geom_histogram(aes(injured_var), alpha = 0.75, fill ="#CFB87C") +
  geom_histogram(aes(uninjured_var), alpha = 0.75) +
  labs(x="Average Variance")
```


```{r Bootstrapping to look at average absolute value running imbalance}
#making a data frame to hold all of the average absolute differences from 0
group_distance <- data.frame(injured_dist = rep(NA,1000),
                uninjured_dist = rep(NA,1000),
                diff_in_dist = rep(NA, 1000))

#bootstrap for variances, 1000 iterations
for(i in 1:1000){
  #random seed
  set.seed(i) 
  
  #taking samples from each of the data sets, same number of rows, replacement true
  injured_sample <- sample_n(injured_data, replace = TRUE, size = 1658)
  uninjured_sample <- sample_n(uninjured_data, replace=TRUE, size = 2405)
  
  #storing the calculated variances in data frame
  group_distance[i,1] = mean(injured_sample$Player.Absolute.Dist)
  group_distance[i,2] = mean(uninjured_sample$Player.Absolute.Dist)
  group_distance[i,3] = group_distance[i,1] - group_distance[i,2]
}
```


```{r Plotting results from averaged boostrapped absolute value running imbalance}
ggplot(data = group_distance, aes(diff_in_dist)) +
  geom_histogram() +
  geom_vline(xintercept = quantile(group_distance$diff_in_dist, 0.05), color= "#CFB87C") +
  geom_vline(xintercept = quantile(group_distance$diff_in_dist, 0.95), color= "#CFB87C") +
  labs(x="Difference in Distance")

ggplot(data = group_distance) +
  geom_histogram(aes(injured_dist), alpha = 0.75, fill ="#CFB87C") +
  geom_histogram(aes(uninjured_dist), alpha = 0.75) +
  labs(x="Average Distance")
```


```{r Removing unnecessary objects from running imbalance variance analysis}
#removing junk that came from the loops
remove(i,team_mean, team_sd, injured_sample, uninjured_sample, group_variances, injured_data, uninjured_data, mean_player_variances, group_distance)
```


## What is a meaningful change? What red flags should go off when we see a week-to-week change in running imbalance?

  Based on the analysis below, there doesn't seem to be any major discernible differences in running imbalance before or following an injury. This suggests that there may not be a direct link between HSI risk and running imbalance directly beforehand. Instead, the trends of many players who were injured seems to have a relatively consistent or trend not entirely dependent on time. 
  Looking at summary statistics of running imbalance in the weeks leading up to and following a hamstring injury, there are also no glaring trends. For this analysis, we looked at the mean and variance in running imbalance per week leading up to and after an injury for all of the injured players with running imbalance data. This showed us that there is no clear indicator of HSI risk in running imbalance or any summary statistic of it. Instead it may be more useful to look at each player's total running imbalance and their individual variance. This seems to be more of a useful tool for differentiating between injured and uninjured athletes.

```{r Calculating weeks before and after injury occurance based on date, how many injuries player has}
#getting running imbalances for just injured players
Injured_Historical_Running <- Historical_Running_clean %>%
  filter(anon_id %in% injured_IDs)

#making new column to represent when in time injury would be, negative means before injury and positive means after injury
Injured_Historical_Running$Weeks.After.Injury <- rep(NA, 1658)
Injured_Historical_Running$Injury.Count <- rep(NA, 1658)

#go through all of the injured players in the data set
for(i in 1:22){
  #get the dates each player was injured
  injury_dates <- unique(Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Date.of.Injury)
  
  for(j in 1:length(injury_dates)){
    #calculate dates for 1, 2, 3, 4 weeks before and after each injury date
    past_1 <- injury_dates[j]-7
    past_2 <- injury_dates[j]-14
    past_3 <- injury_dates[j]-21
    past_4 <- injury_dates[j]-28
    future_1 <- injury_dates[j]+7
    future_2 <- injury_dates[j]+14
    future_3 <- injury_dates[j]+21
    future_4 <- injury_dates[j]+28
    
    #Calculating how many injuries this is
    injury_count <- as.character((length(injury_dates)) - j + 1)
    
    #compare date for each player to date of injury, store in Weeks.After.Injury column, store injury count
    
    #first week after injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > injury_dates[j] & 
                                 Injured_Historical_Running$Date<=future_1,]$Weeks.After.Injury <- "1"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > injury_dates[j] & 
                                 Injured_Historical_Running$Date<=future_1,]$Injury.Count <- injury_count
    
    #second week after injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_1 & 
                                 Injured_Historical_Running$Date<=future_2,]$Weeks.After.Injury <- "2"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_1 & 
                                 Injured_Historical_Running$Date<=future_2,]$Injury.Count <- injury_count
    
    #third week after injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_2 & 
                                 Injured_Historical_Running$Date<=future_3,]$Weeks.After.Injury <- "3"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_2 & 
                                 Injured_Historical_Running$Date<=future_3,]$Injury.Count <- injury_count
    
    #fourth week after injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_3 & 
                                 Injured_Historical_Running$Date<=future_4,]$Weeks.After.Injury <- "4"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_3 & 
                                 Injured_Historical_Running$Date<=future_4,]$Injury.Count <- injury_count    
    #week right before injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < injury_dates[j] & 
                                 Injured_Historical_Running$Date>=past_1,]$Weeks.After.Injury <- "-1"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < injury_dates[j] & 
                                 Injured_Historical_Running$Date>=past_1,]$Injury.Count <- injury_count
    #two weeks before injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_1 & 
                                 Injured_Historical_Running$Date>=past_2,]$Weeks.After.Injury <- "-2"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_1 & 
                                 Injured_Historical_Running$Date>=past_2,]$Injury.Count <- injury_count
    
    #three weeks before injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_2 & 
                                 Injured_Historical_Running$Date>=past_3,]$Weeks.After.Injury <- "-3"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_2 & 
                                 Injured_Historical_Running$Date>=past_3,]$Injury.Count <- injury_count
        
    #four weeks before injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_3 & 
                                 Injured_Historical_Running$Date>=past_4,]$Weeks.After.Injury <- "-4"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_3 & 
                                 Injured_Historical_Running$Date>=past_4,]$Injury.Count <- injury_count
    
    #Date of Injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date == injury_dates[j],]$Weeks.After.Injury <- "0"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date == injury_dates[j],]$Injury.Count <- injury_count
  }
}


#making weeks after injury and injury count into a factor and combining data sets back together
Injured_Historical_Running <- Injured_Historical_Running %>%
  mutate(Weeks.After.Injury = factor(Weeks.After.Injury),
         Injury.Count = factor(Injury.Count))

Historical_Running_clean <- left_join(Historical_Running_clean, Injured_Historical_Running)

#getting rid of junk that was from the loop
remove(future_1, future_2, future_3, future_4, i, injury_count, injury_dates, j, past_1, past_2, past_3, past_4)
remove(Injured_Historical_Running)
```


```{r}
for(i in 1:22){
  p <- ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id==injured_IDs[i],], aes(Date, Running.Imbalance)) +
    geom_line() +
    geom_point(aes(color=Weeks.After.Injury)) +
    geom_vline(xintercept = Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Date.of.Injury) +
    labs(title=injured_IDs[i])+
    scale_color_manual(values = c("-4"="green", "-3"="yellow", "-2"="orange", "-1"="red", "0"="black","1"= "purple", "2"="navy", "3"="blue", "4"="skyblue")) +
    theme_minimal() +
    labs(title="Running Imbalance")
  
  print(p)
}
```


```{r}
#making summary statistics of running imbalance per week relative to injury
Historical_Running_clean <- Historical_Running_clean %>%
  group_by(anon_id, Injury.Count, Weeks.After.Injury) %>%
  mutate(Weeks.After.Injury.Variability = var(Running.Imbalance),
         Weeks.After.Injury.Mean = mean(abs(Running.Imbalance))) %>%
  ungroup()
```


```{r warning=FALSE}
for(i in 1:22){
  #looking at mean running imbalance per week before and after injury for each injured player
  p <- ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id == injured_IDs[i],], aes(Date, Weeks.After.Injury.Mean, group=Injury.Count)) +
  geom_line() +
  geom_point(aes(color=Weeks.After.Injury)) +
    xlim(min(Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Date.of.Injury)-30, max(Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Date.of.Injury)+30) +
  geom_vline(xintercept = Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Date.of.Injury) +
    labs(title=injured_IDs[i]) +
    scale_color_manual(values = c("-4"="green", "-3"="yellow", "-2"="orange", "-1"="red", "0"="black","1"= "purple", "2"="navy", "3"="blue", "4"="skyblue"))
  
  print(p)
}

```


```{r}
for(i in 1:22){
  #looking at variance in running imbalance per week before and after injury for each injured player
  p <- ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id == injured_IDs[i],], aes(Date, Weeks.After.Injury.Variability, group=Injury.Count)) +
  geom_line() +
  geom_point(aes(color=Weeks.After.Injury)) +
    xlim(min(Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Date.of.Injury)-30, max(Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Date.of.Injury)+30) +
  geom_vline(xintercept = Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Date.of.Injury) +
    labs(title=injured_IDs[i]) +
    scale_color_manual(values = c("-4"="green", "-3"="yellow", "-2"="orange", "-1"="red", "0"="black","1"= "purple", "2"="navy", "3"="blue", "4"="skyblue"))
  
  print(p)
}
```


```{r removing unnecessary objects from section}
remove(i, p)
```

## Is running imbalance sensitive enough of a metric to use as a prognosis tool versus a rehab tool?

```{r}
Incident_Report_clean <- Incident_Report_clean %>%
  filter(!is.na(Injury.Prognosis))%>%
  mutate(Expected.Start.Return = as.Date(ifelse(Injury.Prognosis=="No Expected Time Loss",
                                        Date.of.Injury,
                                        ifelse(Injury.Prognosis=="Less than 1 Week",
                                               Date.of.Injury,
                                               ifelse(Injury.Prognosis=="1-4 Weeks",
                                                      Date.of.Injury+days(7),
                                                      Date.of.Injury+days(28))))),
         Expected.End.Return = as.Date(ifelse(Injury.Prognosis=="No Expected Time Loss",
                                        Date.of.Injury,
                                        ifelse(Injury.Prognosis=="Less than 1 Week",
                                               Date.of.Injury+days(7),
                                               ifelse(Injury.Prognosis=="1-4 Weeks",
                                                      Date.of.Injury+days(28),
                                                      Date.of.Injury+days(56)))))) %>%
  group_by(anon_id, Date.of.Injury) %>%
  mutate(Actual.Return = Date.of.Injury+days(sum(Days.in.Status))) %>%
  ungroup()
```


```{r}
for(i in 1:22){
  #looking at mean running imbalance per week before and after injury for each injured player
  p <- ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id == injured_IDs[i],], aes(Date, Running.Imbalance, group=Injury.Count)) +
  geom_line(linewidth=0.1) +
  geom_point() +
  geom_vline(xintercept = Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Date.of.Injury, color="red") +
    geom_vline(xintercept = Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Expected.Start.Return, color="purple", linetype=3) +
    geom_vline(xintercept = Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Expected.End.Return, color="purple", linetype=3) +
    geom_vline(xintercept = Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Expected.End.Return, color="purple", linetype=1) +
    labs(title=injured_IDs[i])
  
  print(p)
}

```


