---
title: "Data Analysis Notebook"
output: html_notebook
---

```{r}
#Authors: [INSERT NAMES HERE]
#Author Date: 
#Purpose: The purpose of this notebook is to house all data set transformation, cleansing, visualization, statistical analysis, and note-taking for the 2025 CU Athletic Department Sports Science Internship Program

#LAST UPDATED: 

#Including helpful libraries
library(tidyverse)
library(readxl)
library(aod)
```


# Data Cleaning
```{r Loading in the data sets}
#loading in the Catapult data to look at sprinting values
Catapult_Session <- read_csv("data-sets/data-sets-uncompressed/data-sets-compressed/Running Imbalance and Speed/Catapult Session - Outdoor FB.csv")

#loading in the Historical Running data to look at running imbalance values
Historical_Running <- read_csv("data-sets/data-sets-uncompressed/data-sets-compressed/Running Imbalance and Speed/Compiled Historical Running Imbalance FB.csv")

#loading in the Incident Report to look at HSIs
Incident_Report <- read_csv("data-sets/data-sets-uncompressed/data-sets-compressed/Running Imbalance and Speed/Incident Report FB IDs.csv")

```

```{r Cleaning Catapult_Session}
Catapult_Session_clean <- Catapult_Session %>%
  #putting the date as a date class
  mutate(Date = as.Date(Date, "%m/%d/%Y")) %>%
  #only selecting important columns for this analysis
  select(anon_id, Date, Age, Primary.Position, Total.Distance, Period.Name, Total.Duration..min., Velocity.Band.1.Total.Distance, Velocity.Band.2.Total.Distance, Velocity.Band.3.Total.Distance, Velocity.Band.4.Total.Distance, Velocity.Band.5.Total.Distance, Velocity.Band.6.Total.Distance, Velocity.Band.7.Total.Distance, Velocity.Band.8.Total.Distance, Maximum.Velocity, Average.Velocity) %>%
  #calculating each player's maximum velocity
  group_by(anon_id) %>%
  mutate(Player.Max.Velocity = max(na.omit(Maximum.Velocity))) %>%
  ungroup() %>%
  #only selecting data from January 1, 2024 and on
  filter(Date >= "2024-01-01")

head(Catapult_Session_clean)
```

```{r Cleaning Historical_Running}
Historical_Running_clean <- Historical_Running %>%
  #taking out rows that don't have data
  filter(Running.Imbalance != "n/a") %>%
  #putting running imbalance as a number and converting the date to a date class
  mutate(Running.Imbalance = as.numeric(Running.Imbalance),
         Date = as.Date(Date, "%m/%d/%Y")) %>%
  #only using data from January 1, 2024 and on
  filter(Date >= "2024-01-01") %>%
  mutate(X=1:4063) %>%
  #making days January 1, 2024
  group_by(anon_id) %>%
  mutate(Days.Since.Start = as.numeric(Date - min(Date))) %>%
  ungroup()

head(Historical_Running_clean)
```

```{r Cleaning Incident_Report}
Incident_Report_clean <- Incident_Report %>%
  #filtering for only hamstring injuries
  filter(OSICS14.Code == "TM1",
         Status != "Full Go")  %>%
  #getting the date of the injury as a date class
  mutate(Date = as.Date(Date, "%m/%d/%Y"),
         Date.of.Injury = as.Date(Date.of.Injury...Onset.of.symptoms, "%m/%d/%Y"),
         Examination.Date = as.Date(Examination.Date, "%m/%d/%Y")) %>%
  #only selecting relevant columns for this analysis
  select(anon_id, Position, Date, Date.of.Injury, Time.of.Injury, Side, OSICS.Injury.Diagnosis, Coach.s.Diagnosis, Recurrence.of.Injury, Choose.Season, Onset.of.Symptoms, Injury.Prognosis, General.Mechanism, Specific.Mechanism, Injured.While., Type.of.Event, Season., Status, Days.in.Status) %>%
  group_by(anon_id, Date.of.Injury) %>%
  mutate(Days.Out = sum(Days.in.Status)) %>%
  ungroup()

head(Incident_Report_clean)
```

```{r Identifying players that have data in both data sets}
#taking the IDs of players who are and aren't injured
all_IDs <- unique(Historical_Running_clean$anon_id)
#taking IDs that were injured and also have running imbalance data
injured_IDs <- intersect(unique(Incident_Report_clean$anon_id), all_IDs)
#taking all players with running imbalance data that don't have an injury
uninjured_IDs <- unique((Historical_Running_clean %>%
  filter(!anon_id %in% injured_IDs))$anon_id)
```

```{r Filtering out players without proper data}
#injured players who also have running imbalance data
Incident_Report_clean <- Incident_Report_clean %>%
  filter(anon_id %in% injured_IDs)

#all players that only have running imbalance data or have both running imbalance data and incidence report
Historical_Running_clean <- Historical_Running_clean %>%
  filter(anon_id %in% injured_IDs | anon_id %in% uninjured_IDs)
```

```{r Removing all unimportant objects created during cleaning}
#removing all_IDs vector and uncleaned data sets
remove(all_IDs)
remove(Incident_Report)
remove(Historical_Running)
remove(Catapult_Session)
```

# Section 1: Running Speed

## How often are athletes reaching ≥ 90% maximum velocity throughout a training season?

## Should we consider the number of sprinting efforts that athletes are completing?

## Are relative efforts and bands more advantageous than the absolute bands provided?

## How does sprinting exposure (# of efforts, % max reached) relate to incidence of hamstring injuries?



# Section 2: Running Imbalance

## What is the variation at the team level and at each individual athlete level?

```{r Looking at team and player variances of running imbalance}
#team variation
Historical_Running_clean %>%
  summarize(Team_Variation = var(Running.Imbalance))

#individual player variation
Historical_Running_clean %>%
  group_by(anon_id) %>%
  summarize(Player_Variation = var(na.omit(Running.Imbalance))) %>%
  ungroup() 
```

```{r Running imbalance measurements for whole team since January 1, 2024}
#calculating mean and variance for team data
team_mean <- mean(Historical_Running_clean$Running.Imbalance)
team_sd <- sd(Historical_Running_clean$Running.Imbalance)

#making scatter plot of team running imbalance data throughout season
ggplot(Historical_Running_clean, aes(Date, Running.Imbalance)) +
  geom_point(alpha = 0.3) +
  geom_hline(yintercept = team_mean, color = "#CFB87C") +
  geom_hline(yintercept = team_mean + team_sd) +
  geom_hline(yintercept = team_mean - team_sd) +
  geom_hline(yintercept = team_mean + (2*team_sd), color = "#A2A4A3") +
  geom_hline(yintercept = team_mean - (2*team_sd), color = "#A2A4A3") +
  labs(title = "Team Running Imbalance Since January 1, 2024", y="Running Imbalance (%)",
       subtitle = "\u03BC = 0.08623412, \u03C3^2 = 14.94215")

#making histogram of team running imbalance data
ggplot(Historical_Running_clean, aes(Running.Imbalance)) +
  geom_histogram() +
  labs(title = "Team Running Imbalance Since January 1, 2024", x="Running Imbalance")
```

```{r Looking at distributions for injured and uninjured separately}
#taking the IDs of players who are and aren't injured
all_IDs <- unique(Historical_Running_clean$anon_id)
#taking IDs that were injured and also have running imbalance data
injured_IDs <- intersect(unique(Incident_Report_clean$anon_id), all_IDs)
#taking all players with running imbalance data that don't have an injury
uninjured_IDs <- unique((Historical_Running_clean %>%
  filter(!anon_id %in% injured_IDs))$anon_id)

#making histogram for running imbalance of all injured athletes
ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,], aes(Running.Imbalance)) +
  geom_histogram(fill = "#CFB87C", alpha = 0.75) +
  #adding in 95% confidence interval
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,]$Running.Imbalance, 0.025), color = "#CFB87C") +
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,]$Running.Imbalance, 0.975), color = "#CFB87C") +
  xlim(-21,21) +
  labs(title = "Running Imbalance for Players with HSI since January 1, 2024")

#making histogram for running imbalance of all uninjured athletes
ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,], aes(Running.Imbalance)) +
  geom_histogram(fill = "black", alpha = 0.75) +
  #adding in 95% confidence interval
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,]$Running.Imbalance, 0.025)) +
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,]$Running.Imbalance, 0.975)) +
  xlim(-21,21) +
  labs(title = "Running Imbalance for Players without HSI since January 1, 2024")

#Plotting injured and uninjured histograms over top one another
ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,], aes(Running.Imbalance)) +
  geom_histogram(alpha = 0.75) +
  #adding in 95% CI for uninjured players
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,]$Running.Imbalance, 0.025)) +
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,]$Running.Imbalance, 0.975)) +
  geom_histogram(data = Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,], aes(Running.Imbalance), fill = "#CFB87C", alpha = 0.75) +
  #adding in 95% CI for injured players
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,]$Running.Imbalance, 0.025), color = "#CFB87C") +
  geom_vline(xintercept = quantile(Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,]$Running.Imbalance, 0.975), color = "#CFB87C") +
  xlim(-21,21) +
  labs(title = "Running Imbalance for Players with and without HSI since January 1, 2024")
```

```{r looking at trends for injured and uninjured players since 1-1-2024}
#making scatter plot of running imbalance data for injured players
ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,], aes(Date, Running.Imbalance)) + 
  geom_point(alpha = 0.3) +
  ylim(-21,21) +
  labs(title = "Running Imbalance for Players with HSI since January 1, 2024")

#making scatter plot of running imbalance for uninjured players
ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,], aes(Date, Running.Imbalance)) + 
  geom_point(alpha = 0.3) +
  ylim(-21,21) +
  labs(title = "Running Imbalance for Players without HSI since January 1, 2024")
```


### Testing to see if significantly different
```{r Bootstrapping variances of data sets to see if they are different between groups}
#Splitting up the data sets into injured and uninjured groups
injured_data <- Historical_Running_clean[Historical_Running_clean$anon_id %in% injured_IDs,] %>%
  group_by(anon_id) %>%
  mutate(Player.Variance = var(Running.Imbalance)) %>%
  ungroup()

uninjured_data <- Historical_Running_clean[Historical_Running_clean$anon_id %in% uninjured_IDs,] %>%
  group_by(anon_id) %>%
  mutate(Player.Variance = var(Running.Imbalance)) %>%
  ungroup()


#making a data frame to hold all of the within group variances
group_variances <- data.frame(injured_var = rep(NA,1000),
                uninjured_var = rep(NA,1000),
                diff_in_var = rep(NA, 1000))

#bootstrap for variances, 1000 iterations
for(i in 1:1000){
  #random seed
  set.seed(i) 
  
  #taking samples from each of the data sets, same number of rows, replacement true
  injured_sample <- sample_n(injured_data, replace = TRUE, size = 1672)
  uninjured_sample <- sample_n(uninjured_data, replace=TRUE, size = 2391)
  
  #storing the calculated variances in data frame
  group_variances[i,1] = var(injured_sample$Running.Imbalance)
  group_variances[i,2] = var(uninjured_sample$Running.Imbalance)
  group_variances[i,3] = group_variances[i,1] - group_variances[i,2]
}

```



```{r Plotting results from bootstrap of variances}
ggplot(data=group_variances, aes(diff_in_var)) +
  geom_histogram() +
  geom_vline(xintercept = quantile(group_variances$diff_in_var, 0.025), color= "#CFB87C") +
  geom_vline(xintercept = quantile(group_variances$diff_in_var, 0.975), color= "#CFB87C")

ggplot(data=group_variances) +
  geom_histogram(aes(injured_var), alpha = 0.75, fill ="#CFB87C") +
  geom_histogram(aes(uninjured_var), alpha = 0.75)
```
```{r Boostrapping average variance between players for injured and uninjured players}
#making a data frame to hold all of the average player variances between groups
mean_player_variances <- data.frame(injured_var = rep(NA,1000),
                uninjured_var = rep(NA,1000),
                diff_in_var = rep(NA, 1000))

for(i in 1:1000){
  #random seed
  set.seed(i) 
  
  #taking samples from each of the data sets, same number of rows, replacement true
  injured_sample <- sample_n(injured_data, replace = TRUE, size = 1672)
  uninjured_sample <- sample_n(uninjured_data, replace=TRUE, size = 2391)
  
  #storing the calculated variances in data frame
  mean_player_variances[i,1] = mean(na.omit(injured_sample$Player.Variance))
  mean_player_variances[i,2] = mean(na.omit(uninjured_sample$Player.Variance))
  mean_player_variances[i,3] = mean_player_variances[i,1] - mean_player_variances[i,2]
}

```

```{r Plotting results from bootstrap of player variances}
ggplot(data = mean_player_variances, aes(diff_in_var)) +
  geom_histogram() +
  geom_vline(xintercept = quantile(mean_player_variances$diff_in_var, 0.025), color= "#CFB87C") +
  geom_vline(xintercept = quantile(mean_player_variances$diff_in_var, 0.975), color= "#CFB87C")

ggplot(data=mean_player_variances) +
  geom_histogram(aes(injured_var), alpha = 0.75, fill ="#CFB87C") +
  geom_histogram(aes(uninjured_var), alpha = 0.75) +
  labs(x="Average Variance")
```

## What is a meaningful change? What red flags should go off when we see a week-to-week change in running imbalance?
```{r Calculating weeks before and after injury occurance based on date}
#getting running imbalances for just injured players
Injured_Historical_Running <- Historical_Running_clean %>%
  filter(anon_id %in% injured_IDs)

#making new column to represent when in time injury would be, negative means before injury and positive means after injury
Injured_Historical_Running$Weeks.After.Injury <- rep(NA, 1658)
Injured_Historical_Running$Injury.Count <- rep(NA, 1658)

#go through all of the injured players in the data set
for(i in 1:22){
  #get the dates each player was injured
  injury_dates <- unique(Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs[i],]$Date.of.Injury)
  
  for(j in 1:length(injury_dates)){
    #calculate dates for 1, 2, 3, 4 weeks before and after each injury date
    past_1 <- injury_dates[j]-7
    past_2 <- injury_dates[j]-14
    past_3 <- injury_dates[j]-21
    past_4 <- injury_dates[j]-28
    future_1 <- injury_dates[j]+7
    future_2 <- injury_dates[j]+14
    future_3 <- injury_dates[j]+21
    future_4 <- injury_dates[j]+28
    
    #Calculating how many injuries this is
    injury_count <- as.character((length(injury_dates)) - j + 1)
    
    #compare date for each player to date of injury, store in Weeks.After.Injury column
    
    #first week after injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > injury_dates[j] & 
                                 Injured_Historical_Running$Date<=future_1,]$Weeks.After.Injury <- "1"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > injury_dates[j] & 
                                 Injured_Historical_Running$Date<=future_1,]$Injury.Count <- injury_count
    
    #second week after injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_1 & 
                                 Injured_Historical_Running$Date<=future_2,]$Weeks.After.Injury <- "2"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_1 & 
                                 Injured_Historical_Running$Date<=future_2,]$Injury.Count <- injury_count
    
    #third week after injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_2 & 
                                 Injured_Historical_Running$Date<=future_3,]$Weeks.After.Injury <- "3"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_2 & 
                                 Injured_Historical_Running$Date<=future_3,]$Injury.Count <- injury_count
    
    #fourth week after injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_3 & 
                                 Injured_Historical_Running$Date<=future_4,]$Weeks.After.Injury <- "4"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date > future_3 & 
                                 Injured_Historical_Running$Date<=future_4,]$Injury.Count <- injury_count    
    #week right before injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < injury_dates[j] & 
                                 Injured_Historical_Running$Date>=past_1,]$Weeks.After.Injury <- "-1"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < injury_dates[j] & 
                                 Injured_Historical_Running$Date>=past_1,]$Injury.Count <- injury_count
    #two weeks before injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_1 & 
                                 Injured_Historical_Running$Date>=past_2,]$Weeks.After.Injury <- "-2"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_1 & 
                                 Injured_Historical_Running$Date>=past_2,]$Injury.Count <- injury_count
    
    #three weeks before injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_2 & 
                                 Injured_Historical_Running$Date>=past_3,]$Weeks.After.Injury <- "-3"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_2 & 
                                 Injured_Historical_Running$Date>=past_3,]$Injury.Count <- injury_count
        
    #four weeks before injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_3 & 
                                 Injured_Historical_Running$Date>=past_4,]$Weeks.After.Injury <- "-4"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date < past_3 & 
                                 Injured_Historical_Running$Date>=past_4,]$Injury.Count <- injury_count
    
    #Date of Injury
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date == injury_dates[j],]$Weeks.After.Injury <- "0"
    
    Injured_Historical_Running[Injured_Historical_Running$anon_id==injured_IDs[i] &
                                 Injured_Historical_Running$Date == injury_dates[j],]$Injury.Count <- injury_count
  }
}


#making weeks after injury into a factor and combining data sets back together
Injured_Historical_Running$Weeks.After.Injury <- as.factor(Injured_Historical_Running$Weeks.After.Injury)
Historical_Running_clean <- left_join(Historical_Running_clean, Injured_Historical_Running)

#getting rid of junk that was from the loop
remove(future_1, future_2, future_3, future_4, i, injury_count, injury_dates, j, past_1, past_2, past_3, past_4)
remove(Injured_Historical_Running)
```


```{r}
for(i in 1:22){
  p <- ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id==injured_IDs[i],], aes(Date, Running.Imbalance)) +
    geom_line() +
    geom_point(aes(color=Weeks.After.Injury)) +
    geom_vline(xintercept = Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs,]$Date.of.Injury) +
    labs(title=injured_IDs[i])
  
  print(p)
}
```


```{r}
Historical_Running_clean <- Historical_Running_clean %>%
  group_by(anon_id, Weeks.After.Injury) %>%
  mutate(Weeks.After.Injury.Variability = var(Running.Imbalance),
         Weeks.After.Injury.Mean = mean(abs(Running.Imbalance))) %>%
  ungroup()
```


```{r}
for(i in 1:22){
  #going through loop
  p <- ggplot(data=Historical_Running_clean[Historical_Running_clean$anon_id==injured_IDs[i],], aes(Date, Weeks.After.Injury.Mean, color=Weeks.After.Injury)) +
    geom_point() +
    geom_vline(xintercept = Incident_Report_clean[Incident_Report_clean$anon_id==injured_IDs,]$Date.of.Injury) +
    labs(title=injured_IDs[i])
  
  print(p)
}

```

## Is running imbalance sensitive enough of a metric to use as a prognosis tool versus a rehab tool?




